### 2024-06-21 output/tables/tab_outcomes_corr.tex

 & \multicolumn{2}{c|}{ZBI vs EQ-5D} & \multicolumn{2}{c|}{ZBI vs EQ-VAS} & \multicolumn{2}{c|}{EQ-5D vs EQ-VAS} \\ \hline
 & Coefficient & P-value & Coefficient & P-value & Coefficient & P-value \\ \hline
Baseline score vs baseline score (N=84)   & -0.198 & 0.069 & -0.210 & 0.054 & 0.191 & 0.080   \\ \hline
9 months score vs 9 months score (N=66)   & -0.382 & 0.002 & -0.273 & 0.028 & 0.527 & 0.000   \\ \hline
Change in score vs change in score (N=66) & -0.296 & 0.017 & -0.082 & 0.519 & 0.386 & 0.001  \\ \hline